\documentclass[fleqn]{article}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{mathexam}
\usepackage{amsmath}

\ExamClass{Physics II}
\ExamName{Review of Physics I}
\ExamHead{\today}

\let\ds\displaystyle

\begin{document}
\ExamInstrBox{
Please show \textbf{all} your work! Write answers in spaces provided. You have 24 minutes to complete this exam.}
\ExamNameLine
\begin{enumerate}
   \item A car is initially travelling at 30 m/s in a westward direction before beginning to accelerate at a rate of 10 m/s$^2$ towards the east.  
      \begin{enumerate}
	 \item What is the velocity of the car after 4 seconds?
	 
	 \answer
	 \item Is the car not moving at any point in time?
	 
	 \answer
	 
	 \item If the car stops accelerating when its speed is 60 m/s, how long did it accelerate for?
	 \answer
      \end{enumerate}
   \item The year is 2075 and Ottawa has decided to open phase 7 of its public transportation rail network --- a bullet train that maxes out at 175 m/s. Assuming the train is originally moving at maximum speed, how far before the station must the train begin slowing down in order to not accelerate  the discomfort point of 35 m/s$^2$?
   \noanswer[2in]
      \newpage
      
   \item You get into a race with your younger sister. Since she's 5 years younger than you, you give her a 30 meter head start. You run at an average speed of 6 m/s and she runs with an average speed of 5 m/s. You reach the finish line in 20 seconds. How far is the finish line from where you started? Did she beat you? 
     \noanswer[3cm]
      \answer
   \item A high impact probe is launched downwards (initial velocity 500 m/s) from a spacecraft hovering 15,000 m above the surface of Mars. Assuming a constant gravitational acceleration of 3.7 m/s$^2$, does the probe reach the required impact velocity of 650 m/s? If not, how high would it need to be launched from in order to achieve this?
     \noanswer[6cm]
      \answer
      \noanswer[0.5cm]
\end{enumerate}
\end{document}


